Open Data, Open Source and Open Standards in chemistry - The Blue Obelisk five years on

### 2011-04-18 paper.tex
Added a second example of between-project code migration
Pasted in Roberts contribution

\author{Charles A Darwin\correspondingauthor$^{1,2}$%
       \email{Charles A Darwin\correspondingauthor - charles@londonzoo.co.uk}%
      \and 
         Egon L Willighagen$^3$%
         \email{Egon L Willighagen - egon.willighagen@ki.se}
     \and 
         Rajarshi Guha$^4$%
         \email{Rajarshi Guha - guhar@mail.nih.gov}
	 \and 
	     Christoph Steinbeck$^6$%
	         \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
     \and 
         Noel M O'Boyle$^5$%
         \email{Noel M O'Boyle - n.oboyle@ucc.ie}
     \and 
         Peter Murray-Rust$^7$%
         \email{Peter Murray-Rust - pm286@cam.ac.uk}
      \and
         Jane E Doe\correspondingauthor$^2$%
         \email{Jane E Doe\correspondingauthor - jane.e.doe@cambridge.co.uk}
      \and
         Dmitry Pavlov$^8$%
         \email{Dmitry Pavlov - dpavlov@ggasoftware.com}
      }

@@

@@

One of the effects of the Blue Obelisk has been to bring developers
together from different Open Source chemistry projects so that they
look for opportunities to collaborate rather than compete, and to
leverage work done by other projects to avoid duplication of effort.
As an example of this, when in March 2008 the Jmol development team
were looking to add support for energy minimisation, rather than
implement a forcefield from scratch they ported the UFF forcefield
implementation from Open Babel to Jmol. This code has allowed Jmol to
support 2D to 3D conversion of structures (through energy
minimisation).

>>>>

\author{Charles A Darwin\correspondingauthor$^{1,2}$%
       \email{Charles A Darwin\correspondingauthor - charles@londonzoo.co.uk}%
      \and 
         Egon L Willighagen$^3$%
         \email{Egon L Willighagen - egon.willighagen@ki.se}
     \and 
         Rajarshi Guha$^4$%
         \email{Rajarshi Guha - guhar@mail.nih.gov}
	 \and 
	     Christoph Steinbeck$^6$%
	         \email{Christoph Steinbeck - steinbeck@ebi.ac.uk}
     \and 
         Noel M O'Boyle$^5$%
         \email{Noel M O'Boyle - n.oboyle@ucc.ie}
     \and 
         Peter Murray-Rust$^7$%
         \email{Peter Murray-Rust - pm286@cam.ac.uk}
     \and
         Robert Hanson$^8$%
         \email{Robert Hanson - hansonr@stolaf.edu}
      \and
         Jane E Doe\correspondingauthor$^2$%
         \email{Jane E Doe\correspondingauthor - jane.e.doe@cambridge.co.uk}
      \and
         Dmitry Pavlov$^8$%
         \email{Dmitry Pavlov - dpavlov@ggasoftware.com}
      }

@@

Some project have also been adopted by other fields, including biology
and crystallography. A good example
here is Jmol, which has, in the past five years, gone through two major revisions
as part of an accelerated development focus. Starting with Jmol 10.2, released in
April of 2006, Jmol 11.0 added over 100 new features and was released in March of
2007. Three years later, Jmol 12.0 followed with over 700 new features.
Based on the outstanding work of Michael (Miguel) Howard and others, this development
path took Jmol from a "Rasmol/Chime replacement" to a fully fledged molecular
visualization package, including full support for crystallography~\cite{Hanson2010},
efficient delivery of surface data using fast solvent-excluded surface
generation and the highly compressed JVXL format [http://chemapps.stolaf.edu/jmol/docs/misc/JVXL-format.pdf],
 display of molecular orbitals from standard basis set/coefficient data,
the inclusion of dynamic minimization using the UFF force field, and
a full implementation of Daylight SMILES and SMART, with extensions to
conformational and biomolecular substructure searching in the form of Jmol
BioSMARTS.[http://jmol.svn.sourceforge.net/viewvc/jmol/trunk/Jmol/src/org/jmol/smiles/package.html]
Behind the scenes, Jmol has been essentially completely rewritten. A full
mathematical expression evaluator was added, and the scripting language
was folded into a versatile JavaScript-like command language, including a
complete set of program flow commands, several new data types, and
user-defined variables and functions. An exciting non-chemical application
of Jmol has been its incorporation into the open source Sage
Mathematics Project [http://www.sagemath.org/], where Jmol is being used
to deliver three-dimensional depiction of mathematical functions.

@@

One of the effects of the Blue Obelisk has been to bring developers
together from different Open Source chemistry projects so that they
look for opportunities to collaborate rather than compete, and to
leverage work done by other projects to avoid duplication of effort.
As an example of this, when in March 2008 the Jmol development team
were looking to add support for energy minimisation, rather than
implement a forcefield from scratch they ported the UFF forcefield
implementation from Open Babel to Jmol. This code has allowed Jmol to
support 2D to 3D conversion of structures (through energy
minimisation). Similarly, efficient Jmol code for atom-atom rebonding
has been ported to the CDK.
